---
title: "Scenario Comparison"
author: "Thomas Estabrook"
date: "`r Sys.Date()`"
output: html_document
params:
  scenario_column: "Mgt_scenario"
  scenario_column2: "Climate"
  reference_scenario1: "BAU wildfire"
  reference_scenario2: "Base climate"
  color_column: "Color"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.width = 10,
                      fig.height = 7)
```

```{r}
sims.plot.df <- sims.df |>  # add the information from scenario_map
  mutate(Scenario = get(params$scenario_column),
         Scenario2 = get(params$scenario_column2)) |>  
  mutate(Scenario = factor(Scenario, levels=unique(Scenario)),
         Scenario2 = factor(Scenario2, levels=unique(Scenario2)))

scenario_colors <- iwanthue(length(unique(sims.plot.df$Scenario)), 0, 360, 36, 180, 13, 73)
names(scenario_colors) <- unique(sims.plot.df$Scenario)

if(params$scenario_column2 == "Climate"){
  scenario_colors2 <- climate_colors
} else {
  scenario_colors2 <- iwanthue(length(unique(sims.plot.df$Scenario2)))
  names(scenario_colors2) <- unique(sims.plot.df$Scenario2)
}

nsc1 = length(unique(sims.plot.df$Scenario))
nsc2 = length(unique(sims.plot.df$Scenario2))

biomass.all.df <- lapply(file.path(sims.df$Run,'Biomass_Annual_Dynamics.csv'), read_and_label) |> bind_rows() |>
  left_join(sims.plot.df)

fireevents.all.df <- lapply(file.path(sims.df$Run,'scrapple-events-log.csv'), read_and_label) |> bind_rows() |>
  left_join(sims.plot.df)

fires.all.df <- lapply(file.path(sims.df$Run,'Fire_severity_df.csv'), read_and_label) |> bind_rows() |>
  left_join(sims.plot.df) |>
  pivot_longer(c(PercentUnburned, PercentLow, PercentModerate, PercentHigh), names_to = "Severity", names_prefix = "Percent", values_to = "Percent") |>
  mutate(Area_ha = FIRE_SIZE * (Percent/100),
         Severity = ifelse(Severity == "Unburned", "Low", Severity)) |>
  mutate(Severity = factor(Severity, levels = c("Low", "Moderate", "High", labels = c("Low Severity", "Moderate Severity", "High Severity"))))

dnbr.all.df <- lapply(file.path(sims.df$Run,'Fire_all_dnbr_values.csv'), read_and_label) |> bind_rows() |> 
  rename(dNBR = x) |>
  bind_rows(emp.fire.dnbr |> filter(Source == 'MTBS') |> mutate(Run = Source) |> select(dNBR, Run)) |>
  left_join(sims.plot.df |> bind_rows(data.frame("Run" = "MTBS", "Scenario" = "MTBS", "Scenario2" = "Historical")))
```



## Landis Scenario Comparison

### Overview

*Simulation folder:* `landisOutputDir`

*Scenario list:*

```{r}
knitr::kable(sims.plot.df|>group_by(Scenario, Scenario2)|>tally())
```


### Wildfire Comparisons

*Reference Scenario:* `r params$reference_scenario1` with `r params$reference_scenario2`

```{r, fig.cap="Wildfire area burned per severity class (deviation from reference scenario)"}
df <- fires.all.df |>
  filter(IgnitionType!="Prescribed") |>
  group_by(Run, SimulationYear, Scenario, Scenario2, Severity) |>
  summarise(Area_total = sum(Area_ha)) |>  # total annual area in each severity class
  ungroup() |>
  group_by(Run, Scenario, Scenario2, Severity) |>
  mutate(Area_cumulative = cumsum(Area_total)/404.6) |>  # cumulative area for each year to 1000s of acres
  ungroup() |> 
  group_by(Scenario, Scenario2, SimulationYear, Severity) |>
  summarise(Max = max(Area_cumulative), Min = min(Area_cumulative), Mean = mean(Area_cumulative)) |>
  compute_deviation(Sc = params$reference_scenario1, Sc2 = params$reference_scenario2)

ggplot(data = df, aes(x = SimulationYear, y = Mean, color = Scenario, fill = Scenario)) +
  geom_line() +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_color_manual(values = scenario_colors) + scale_fill_manual(values = scenario_colors) +
  facet_grid(Scenario2~Severity) +
  ylab("Area (1000s of Acres)")
```




```{r, fig.cap="Wildfire area burned per severity class (deviation from reference scenario)"}
ggplot(data = df, aes(x = SimulationYear, y = Mean, color = Scenario2, fill = Scenario2)) +
  geom_line() +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_color_manual(values = scenario_colors2) + scale_fill_manual(values = scenario_colors2) +
  facet_grid(Severity~Scenario)
```



```{r, fig.cap="Area (including Rx) burned per severity class (deviation from reference scenario)"}
df <- fires.all.df |>
  group_by(Run, SimulationYear, Scenario, Scenario2, Severity) |>
  summarise(Area_total = sum(Area_ha)) |>  # total annual area in each severity class
  ungroup() |>
  group_by(Run, Scenario, Scenario2, Severity) |>
  mutate(Area_cumulative = cumsum(Area_total)/404.6) |>  # cumulative area for each year to 1000s of acres
  ungroup() |> 
  group_by(Scenario, Scenario2, SimulationYear, Severity) |>
  summarise(Max = max(Area_cumulative), Min = min(Area_cumulative), Mean = mean(Area_cumulative)) |>
  compute_deviation(Sc = params$reference_scenario1, Sc2 = params$reference_scenario2)

ggplot(data = df, aes(x = SimulationYear, y = Mean, color = Scenario, fill = Scenario)) +
  geom_line() +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_color_manual(values = scenario_colors) + scale_fill_manual(values = scenario_colors) +
  facet_grid(Scenario2~Severity) +
  ylab("Area (1000s of Acres)")
```


Area burned per severity class deviation from reference scenario (`r params$reference_scenario1` with `r params$reference_scenario2`):


```{r, fig.cap="Area (including Rx) burned per severity class (deviation from reference scenario)"}
ggplot(data = df, aes(x = SimulationYear, y = Mean, color = Scenario2, fill = Scenario2)) +
  geom_line() +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_color_manual(values = scenario_colors2) + scale_fill_manual(values = scenario_colors2) +
  facet_grid(Severity~Scenario)
```




```{r, fig.cap="Fire severity (pseudo-dNBR) distribution"}
ggplot(data = dnbr.all.df, aes(x = dNBR, fill = Scenario2)) +
  geom_density(alpha = 0.25) + 
  scale_fill_manual(values = scenario_colors2) +
  facet_wrap(~Scenario, ncol=1)
```




```{r, fig.cap="Cohort mortality % distributions"}

```


```{r, fig.cap="Fine Fuel Loading"}
df <- biomass.all.df |>
  filter(Species == "TotalBiomass", 
         PWG %in% c(20, 30, 40, 50)) |>
  group_by(Year, PWG_Name, Scenario, Scenario2) |>
  summarize(
    Max = max(FineFuels_mean_MgHaEco),
    Min = min(FineFuels_mean_MgHaEco),
    Mean = mean(FineFuels_mean_MgHaEco)
  )

ggplot(data = df, aes(x = Year, fill = Scenario, color = Scenario, fill = Scenario)) + 
  geom_line(aes(y = Mean)) +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_fill_manual(values = scenario_colors) + scale_color_manual(values = scenario_colors) +
  facet_grid(Scenario2~PWG_Name) +
  xlab("Simulation Year") + ylab(expression(paste('Fine Fuels (Mg  ',ha^-1,')')))
```



### Biomass Comparisons


```{r, fig.cap="Biomass density"}
df <- biomass.all.df |>
  filter(Species == "TotalBiomass",
         PWG %in% c(20, 30, 40, 50)) |>
  group_by(Year, PWG_Name, Scenario, Scenario2) |>
  summarize(
    Max = max(Biomass_mean_MgHaEco),
    Min = min(Biomass_mean_MgHaEco),
    Mean = mean(Biomass_mean_MgHaEco)
  )

ggplot(data = df, aes(x = Year, fill = Scenario, color = Scenario, fill = Scenario)) + 
  geom_line(aes(y = Mean)) +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_fill_manual(values = scenario_colors) + scale_color_manual(values = scenario_colors) +
  facet_grid(Scenario2~PWG_Name) +
  xlab("Simulation Year") + ylab(expression(paste('Biomass (Mg  ',ha^-1,')')))
```




```{r, fig.cap="Total biomass"}
df <- biomass.all.df |>
  filter(Species == "TotalBiomass",
         PWG %in% c(20, 30, 40, 50)) |>
  group_by(Year, PWG_Name, Scenario, Scenario2) |>
  summarise(
    Mean = mean(Biomass_sum_Mg),
    Max = max(Biomass_sum_Mg),
    Min = min(Biomass_sum_Mg)
  )

ggplot(data = df, aes(x = Year, fill = Scenario, color = Scenario, fill = Scenario)) + 
  geom_line(aes(y = Mean)) +
  geom_ribbon(aes(ymin=Min,ymax=Max),colour=NA,alpha=0.2,size=0.25) +
  scale_fill_manual(values = scenario_colors) + scale_color_manual(values = scenario_colors) +
  facet_grid(Scenario2~PWG_Name) +
  xlab("Simulation Year") + ylab(expression(paste('Biomass (Mg)')))
```

Total biomass deviating from reference scenario (`r params$reference_scenario1` with `r params$reference_scenario2`):

```{r, fig.cap="Total biomass deviation from reference scenario"}
df <- biomass.all.df |>
  filter(Species == "TotalBiomass",
         PWG %in% c(20, 30, 40, 50)) |>
  group_by(Year, PWG_Name, Scenario, Scenario2) |>
  summarise(
    Mean = mean(Biomass_sum_Mg),
    Max = max(Biomass_sum_Mg),
    Min = min(Biomass_sum_Mg)
  ) |>
  compute_deviation(Sc = params$reference_scenario1, Sc2 = params$reference_scenario2)

ggplot(data = df, aes(x = Year, fill = Scenario, color = Scenario, fill = Scenario)) + 
  geom_line(aes(y = Mean)) +
  geom_ribbon(aes(ymin=Min,ymax=Max),colour=NA,alpha=0.2,size=0.25) +
  scale_fill_manual(values = scenario_colors) + scale_color_manual(values = scenario_colors) +
  facet_grid(Scenario2~PWG_Name) +
  xlab("Simulation Year") + ylab(expression(paste('Biomass (Mg)')))
```


```{r, fig.cap="Dead wood pool"}
df <- biomass.all.df |>
  filter(Species == "TotalBiomass", 
         PWG %in% c(20, 30, 40, 50)) |>
  group_by(Year, PWG_Name, Scenario, Scenario2) |>
  summarize(
    Max = max(DeadWood_mean_MgHaEco),
    Min = min(DeadWood_mean_MgHaEco),
    Mean = mean(DeadWood_mean_MgHaEco)
  ) |>
  filter(!is.na(Mean))

ggplot(data = df, aes(x = Year, fill = Scenario, color = Scenario, fill = Scenario)) + 
  geom_line(aes(y = Mean)) +
  geom_ribbon(aes(ymin=Min,ymax=Max),alpha=0.2, linewidth = NA) +
  scale_fill_manual(values = scenario_colors) + scale_color_manual(values = scenario_colors) +
  facet_grid(Scenario2~PWG_Name) +
  xlab("Simulation Year") + ylab(expression(paste('Dead Wood (Mg  ',ha^-1,')')))
```


TODO: Biomass by functional type:




### Harvest Comparisons

```{r, fig.cap="Total harvested biomass by PWG"}
df <- biomass.all.df |>
  filter(Species == "TotalBiomass",
         PWG %in% c(20, 30, 40, 50),
         Harvest_outputs == "yes",
         Year > 0) |>
  group_by(Year, PWG_Name, Scenario, Scenario2) |>
  summarise(
    Mean = mean(Harvested_Mg),
    Max = max(Harvested_Mg),
    Min = min(Harvested_Mg)
  )

if(nrow(df) > 0){
  ggplot(data = df, aes(x = Year, fill = Scenario, color = Scenario, fill = Scenario)) + 
    geom_line(aes(y = Mean)) +
    geom_ribbon(aes(ymin=Min,ymax=Max),colour=NA,alpha=0.2,size=0.25) +
    scale_fill_manual(values = scenario_colors) + scale_color_manual(values = scenario_colors) +
    facet_grid(Scenario2~PWG_Name) +
    xlab("Simulation Year") + ylab(expression(paste('Harvested Biomass (Mg)')))
}

```

```{r, fig.cap="Total harvested biomass by PWG"}
if(nrow(df) > 0){
  ggplot(data = df, aes(x = Year, fill = Scenario2, color = Scenario2, fill = Scenario2)) + 
    geom_line(aes(y = Mean)) +
    geom_ribbon(aes(ymin=Min,ymax=Max),colour=NA,alpha=0.2,size=0.25) +
    scale_fill_manual(values = scenario_colors2) + scale_color_manual(values = scenario_colors2) +
    facet_grid(Scenario~PWG_Name) +
    xlab("Simulation Year") + ylab(expression(paste('Harvested Biomass (Mg)')))
}

```















